---
title: "Example-03"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Example-03}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Load parameters and initial conditions from files
This example shows how to load parameters and initial conditions.    

The tutorial is adapted from Visual Basic scripts provided by Javier Fluixa Sanmartin (CREALP). 

## Prerequisites
Install R, RTools, RStudio, RSMinerve and the R package rClr (see [README](https://github.com/hydrosolutions/RSMinerveR#readme) for instructions and links).   

Copy this vignette to your local computer and adapt the paths.   
  
Open the model in \code{RSMinerveR/data/Model.rsm} and load & link the data base with the model objects. If you don't know how to do this please follow the examples in the RS Minerve User Manual prior to continuing with this tutorial.  

Open the ```Selection and plots``` tab by clicking on its icon in the in the modules toolbar. Define a selection of model results to visualize (here we define two result groups: ```DischargeGlacier``` and ```SnowWaterEquivalent```. Save the model.  


```{r setup, message=FALSE,warning=FALSE,results='hide'}
library(RSMinerveR)
library(rClr)
library(tidyverse)
library(lubridate)
library(R.utils)
```


## Loading Assemblies

```{r load-assemblies}
# Adapt path in dir_RSM_install to your local installation of RSMinerve. 
dir_RSM_install <- "C:/Program Files (x86)/RS MINERVE"

clrLoadAssembly(file.path(dir_RSM_install, 'log4net.dll'))
clrLoadAssembly(file.path(dir_RSM_install, 'Microsoft.Practices.Prism.Mvvm.dll'))
clrLoadAssembly(file.path(dir_RSM_install, 'RSMinerve.RS.dll'))
clrLoadAssembly(file.path(dir_RSM_install, 'RSMinerve.DB.dll'))
clrLoadAssembly(file.path(dir_RSM_install, 'RSMinerve.Base.dll'))
```

## Model settings

```{r model-settings}
model_file <- if (file.exists("../inst/extdata/Tutorial_Model.rsm")) {
   "../inst/extdata/Tutorial_Model.rsm"
} else {
  "https://raw.githubusercontent.com/hydrosolutions/RSMinerveR/blob/main/inst/extdata/Tutorial_Model.rsm"
}
input_dataset <- if (file.exists("../inst/extdata/Tutorial_DataMeteo.dsx")) {
   "../inst/extdata/Tutorial_DataMeteo.dsx"
} else {
  "https://raw.githubusercontent.com/hydrosolutions/RSMinerveR/blob/main/inst/extdata/Tutorial_DataMeteo.dst"
}
parameter_file <- if (file.exists("../inst/extdata/Tutorial_Parameters.txt")) {
   "../inst/extdata/Tutorial_Parameters.txt"
} else {
  "https://raw.githubusercontent.com/hydrosolutions/RSMinerveR/blob/main/inst/extdata/Tutorial_Parameters.txt"
}
ic_file <- if (file.exists("../inst/extdata/Tutorial_IC.txt")) {
   "../inst/extdata/Tutorial_IC.txt"
} else {
   "https://raw.githubusercontent.com/hydrosolutions/RSMinerveR/blob/main/inst/extdata/Tutorial_IC.txt"
}


saveDataInDstFile <-  TRUE

# The RS Minerve documentation describes the date format to be "%d.%m.%Y" 
start_date <- "02.09.2013 00:00:00"  # format = "%d.%m.%Y %H:%M:%S"
end_date <- "09.09.2013 00:00:00"  # format = "%d.%m.%Y %H:%M:%S"

simulationTimeStep <- "600"  
recordingTimeStep <- "3600"
timeStepUnit <- "Seconds"

```


The paths for the result files need to be full paths. Below are example paths. You'll need to adapt these. 

```{r, echo=FALSE}
output_path <- file.path("C:", "Users", "marti", "Dropbox (hydrosolutions)", 
                         "HYDRO4U_TECHNICAL_WORK", "MODELS", 
                         "test_RSMinerveRvignettes")

if (dir.exists(output_path)) {
  
  results_GlacierDischarge_savePath <- file.path(output_path, 
                                                 "tutorial03-results_GlacierDischarge.dsx")
  results_SWE_savePath <- file.path(output_path, "tutorial03-results_SWE.dsx")
  presimreport_path <- file.path(output_path, "tutorial03-preSimulationReport.txt")
  postsimreport_path <- file.path(output_path, "tutorial03-postSimulationReport.txt")
  ic_save_folder <- output_path
  
} else {
  
  cat("Warning: Result folder not found. Attemping to write to temporary folder in the current working directory. \n")
  dir.create("temporary_RSMinerveR_output")
  
  if (dir.exists("temporary_RSMinerveR_output")) {
    results_GlacierDischarge_savePath <- normalizePath(file.path("temporary_RSMinerveR_output", 
                                                "tutorial03-results_GlacierDischarge.dsx"), 
                                      mustWork = FALSE)
    results_SWE_savePath <- normalizePath(file.path("temporary_RSMinerveR_output", 
                                                    "tutorial03-results_SWE.dsx"), 
                                          mustWork = FALSE)
    presimreport_path <- normalizePath(file.path("temporary_RSMinerveR_output",
                                                 "tutorial03-preSimulationReport.txt"), 
                                       mustWork = FALSE)
    postsimreport_path <- normalizePath(file.path("temporary_RSMinerveR_output", 
                                                  "tutorial03-postSimulationReport.txt"), 
                                       mustWork = FALSE)
    ic_save_folder <- normalizePath("temporary_RSMinerveR_output", 
                                    mustWork = FALSE)
  }

}
```

```{r result-paths, eval=FALSE}
results_GlacierDischarge_savePath <- 
  "C:/Users/<username>/Documents/scriping_rsm/tutorial03-results_GlacierDischarge.dsx"
results_SWE_savePath <- 
  "C:/Users/<username>/Documents/scriping_rsm/tutorial03-results_SWE.dsx"
presimreport_path <- 
  "C:/Users/<username>/Documents/scriping_rsm/tutorial03-preSimuReport.txt"
postsimreport_path <- 
  "C:/Users/<username>/Documents/scriping_rsm/tutorial03-postSimuReport.txt"
ic_save_folder <- 
  "C:/Users/<username>/Documents/scriping_rsm"
```


## Applying settings and running the model
The \code{clrCall} commands allow to call the Visual Basics commands documented in the RS Minerve Technical Manual. The calls return NULL when successful. Check your local paths to see if the output files have been created.  
```{r run-model}
# Define a clr task
rsm <- clrNew("RSMinerve.RS.Task")  

# Start-up the model
clrCall(rsm, "Start", model_file)

# Load data
clrCall(rsm, "LoadDatasetAndSetDates", input_dataset, TRUE, TRUE)

# Set simulation parameters
clrCall(rsm, "SetDates", start_date, end_date)
# SetSimulationTimeStep and SetRecordingTimeStep require characters
clrCall(rsm, "SetSimulationTimeStep", simulationTimeStep, timeStepUnit) 
clrCall(rsm, "SetRecordingTimeStep", recordingTimeStep, timeStepUnit)  

# Load initial conditions from file and copy file
clrCall(rsm, "LoadInitialConditionsFromFileAndCopyFile", ic_file, ic_save_folder)

# Load parameters from file
clrCall(rsm, "LoadParametersFromFile", parameter_file)

# Write a pre-simulation report to check for possible errors (requires full 
# path)
clrCall(rsm, "SavePreSimulationReportAs", presimreport_path)  

# Run a simulation
clrCall(rsm, "Simulate")

# Write a post-simulation report to check for possible errors (requires full 
# path)
clrCall(rsm, "SavePostSimulationReportAs", postsimreport_path)  

# Save selected results (requires full path)
clrCall(rsm, "SaveSelectionResultsAs", results_GlacierDischarge_savePath, 
        saveDataInDstFile, "DischargeGlacier")  
clrCall(rsm, "SaveSelectionResultsAs", results_SWE_savePath, 
        saveDataInDstFile, "SnowWaterEquivalent")  

# Stop task
clrCall(rsm, "Stop")
```

## Read results
```{r, message=FALSE, warning=FALSE, results='hide'}
# Calculate the number of time steps to read for each model component, i.e. the 
# chunk size. Including the header of each chunk. 
chunk_size <- getChunkSize(
  lubridate::as_datetime(start_date,format = "%d.%m.%Y %H:%M:%S"), 
  lubridate::as_datetime(end_date, format = "%d.%m.%Y %H:%M:%S"), 
  recordingTimeStep
) 
result_glacier <- readResultDST(results_GlacierDischarge_savePath, chunk_size)
result_swe <- readResultDST(results_SWE_savePath, chunk_size)

```


## Visualisation example

### Glacier melt contributions to discharge
```{r, fig.fullwidth=TRUE, fig.width=6}
result_glacier |> 
ggplot() + 
  geom_point(aes(Datetime, Value, colour = Object)) +
  theme_bw()
```

### Snow water equivalent in non-glaciated sub-basins
```{r, fig.fullwidth=TRUE, fig.width=6}
result_swe |> 
ggplot() + 
  geom_point(aes(Datetime, Value, colour = Object)) +
  theme_bw()
```
